A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-07 Paper/main.tex
update

\textbf{Multi-domain Buffer Overflows attacks.}
According to~\cite{piromsopa2006buffer}, a buffer-overflow attack is an attack that (possibly implicitly) uses memory-manipulating operations to overflow a buffer which results in the modification of an address to point to malicious or unexpected code. In this case, we will integrate data from various log sources; to be more exact, we will examine both network and web server access logs to determine when a server buffer overflow attack has been performed. We search through network packets for binary data; then web server's access logs are examined to see if a corresponding event, in which the web server correctly terminates the connection after receiving some binary data, can be located. If no matching log record is discovered within a predetermined timeout, it means that the buffer overflow attack was successful. The following formula captures this scenario:
\begin{align*}
    & \underline{\text{min}}\ \mathtt{EventuallyClosed}(\underline{\text{long}}\ t,\underline{\text{long}}\ d,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2) = (time-t<d)\\
    & \quad\quad\land ((ip1=i1\land ip2=i2\land log=web\land type=closed)\lor \bigcirc\mathtt{EventuallyClosed}(t,d,i1,i2) ) \\
    & \underline{\text{mon}}\ \mathtt{BufferSafe} =\ \mathtt{Always}((log=network\land type=binary) \\
    & \quad\quad\rightarrow\mathtt{EventuallyClosed}(time,100,ip1,ip2))
\end{align*}
In each record we have the $log$ field which indicates the source of the log, the $time$ represents the istant in which the event occurs, $ip1$ represents the source address, $ip2$ represents the destination IP address and finally the $type$ field can be $binary, closed,\ldots$ We assume that a connection is identified by the tuple $(ip1,ip2)$. In order, the $\mathtt{EventuallyClosed}$ rule states that the connection between IP addresses $i1$ and $i2$ must end eventually within time $d$; finally, the monitor rule $\mathtt{BufferSafe}$ states that it is always the case that if there is an event of $binary$ access in the network log then eventually within time $100$ there must be a matching event in the webserver log that denotes the closing of the connection.

\textbf{Password guessing attacks.}
\section{Conclusions}

>>>>

\textbf{Multi-domain Buffer Overflows attacks.}
According to~\cite{piromsopa2006buffer}, a buffer-overflow attack is an attack that (possibly implicitly) uses memory-manipulating operations to overflow a buffer which results in the modification of an address to point to malicious or unexpected code. In this case, we will integrate data from various log sources; to be more exact, we will examine both network and web server access logs to determine when a server buffer overflow attack has been performed. We search through network packets for binary data; then web server's access logs are examined to see if a corresponding event, in which the web server correctly terminates the connection after receiving some binary data, can be located. If no matching log record is discovered within a predetermined timeout, it means that the buffer overflow attack was successful. The following formulas captures this scenario:
\begin{align*}
    & \underline{\text{min}}\ \mathtt{EventuallyClosed}(\underline{\text{long}}\ t,\underline{\text{long}}\ d,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2) = (time-t<d)\\
    & \quad\quad\land ((ip1=i1\land ip2=i2\land log=web\land type=closed)\lor \bigcirc\mathtt{EventuallyClosed}(t,d,i1,i2) ) \\
    & \underline{\text{mon}}\ \mathtt{BufferSafe} =\ \mathtt{Always}((log=network\land type=binary) \\
    & \quad\quad\rightarrow\mathtt{EventuallyClosed}(time,100,ip1,ip2))
\end{align*}
In each record we have the $log$ field which indicates the source of the log, the $time$ represents the istant in which the event occurs, $ip1$ represents the source address, $ip2$ represents the destination IP address and finally the $type$ field can be $binary, closed,\ldots$ We assume that a connection is identified by the tuple $(ip1,ip2)$. In order, the $\mathtt{EventuallyClosed}$ rule states that the connection between IP addresses $i1$ and $i2$ must end eventually within time $d$; finally, the monitor rule $\mathtt{BufferSafe}$ states that it is always the case that if there is an event of $binary$ access in the network log then eventually within time $100$ there must be a matching event in the webserver log that denotes the closing of the connection.

\textbf{Password guessing attacks.}
Assuming the system accepts telnet connections to remote login; it won't take long until the password is compromised if an attacker is permitted to guess any amount of passwords for a specific user name. If a user guesses an incorrect password more than three times in a short amount of time, most systems end the telnet session. To prevent an attacker from succeeding by starting numerous temporary sessions, some systems cap the total number of unsuccessful login attempts over a longer time frame; in this case we need to have access to the host's audit logs. The formulas that capture this scenario are the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{Failure}() = (type=login)\land \neg success\\
    & \underline{\text{max}}\ \mathtt{Guess}(\underline{\text{string}}\ i) = (ip=i)\land \mathtt{Failure}()\\
    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i,\underline{\text{int}}\ C) = (time-t<d)\\
    & \quad\quad\rightarrow((\mathtt{Guess}(i)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i,C)))  \\
    & \quad\quad\land(\neg \mathtt{Guess}(i)\rightarrow\mathtt{Counter}(t,d,c,i,C))) \\
    & \underline{\text{mon}}\ \mathtt{PassGuessSafe} =\ \mathtt{Always}(\mathtt{Failure}()\rightarrow\mathtt{Counter}(time,300,1,ip,3)) 
\end{align*}
as we can see, the rules here are numerous. In $\mathtt{Counter}$, the arguments $t, d, c, i, C$ represent the rule invocation time, the timeout period, the current number of unsuccessful-guesses count, the source IP address doing the guess, and the threshold count. When the number of guesses from the IP address $i$ exceeds the threshold count $C$ within the timeout period $d$, an attack has taken place. The parameterized rule $\mathtt{Counter}$ starts with the initial count $c$ set to 1 whenever there is a $\mathtt{Failure}$ during login. The number of guesses is increased by one for each instance of an event signifying a failed login attempt from the same IP address within the timeout period $d$. The rule $\mathtt{Counter}$ also determines whether the local counter $c$ surpasses the threshold $C$ within time d, and if it does, the entire rule become false, indicating an attack. Finally, with the rule $\mathtt{PassGuessSafe}$ we assure that whenever there is a failed login from a specific IP address, we make sure that within time $300$ seconds the number of login-failures from the same IP address must be less than or equal to $3$.

\section{Conclusions}
